Why aren't significance tests commonly used for linear regression diagnostics?

### 2022-09-20 paper_comparison/paper_comparison/paper_comparison.tex
omit the analysis section for faster knitting.

With our knowledge, the effectiveness of visual test relative to the
equivalent conventional test in regression diagnostics has not been
validated. Particularly, its ability to detect non-linearity and
heteroskedasticity compared to model structural test and Breusch--Pagan
test. Meanwhile, as mentioned in the \ref{introduction}
\protect\hyperlink{introduction}{introduction} section

For the purpose of investigating the differences between conventional
hypothesis testing and graphical diagnostics in the application of
linear regression diagnostics, three experiments were conducted.

The experiment I has ideal scenario for conventional testing, where the
visual test is not expected to outperform the conventional test. The
experiment II is a scenario where the conventional test is an
approximate test, in which the visual test may have a chance to match
the performance of the conventional test.

Subjects for both experiments were recruited from an crowdsourcing
platform called Prolific {[}ref here{]}. Prescreening procedure was
applied during the recruitment, subjects were required to be fluent in
English, with \(98\%\) minimum approval rate in other studies and 10
minimum submissions.

@@

\hypertarget{nonlinearity}{%
\subsubsection{Nonlinearity}\label{nonlinearity}}

\hypertarget{old-stuff}{%
\section{Old stuff}\label{old-stuff}}

\hypertarget{computing-power-curves}{%
\subsection{Computing power curves}\label{computing-power-curves}}

For the purpose of examining the effectiveness of visual test in
regression diagnostics, two experiments were conducted. The experiment I
has ideal scenario for conventional testing, where the visual test is
not expected to outperform the conventional test. The experiment II is a
scenario where the conventional test is an approximate test, in which
the visual test may have a chance to match the performance of the
conventional test.

Subjects for both experiments were recruited from an crowdsourcing
platform called Prolific {[}ref here{]}. Prescreening procedure was
applied during the recruitment, subjects were required to be fluent in
English, with \(98\%\) minimum approval rate in other studies and 10
minimum submissions.

A pool of 12 different lineups with obvious visual patterns were
generated for experiment I and experiment II respectively. In every
block of 20 lineups that presented to a subject, two out of 12 lineups
were included as attention checks. A subject's submission was only
accepted if the actual data plot was identified for at least one
attention check. Data of rejected submissions were discarded
automatically to maintain the overall data quality.

\hypertarget{non-linearity}{%
\subsection{Non-linearity}\label{non-linearity}}

@@

Model misspecification presents since the null model leaves out the
higher order term.

@@

\hypertarget{heteroskedasticity-1}{%
\subsection{Heteroskedasticity}\label{heteroskedasticity-1}}

Experiment II is designed to study the ability of human subjects to
detect the appearance of a heteroskedasticity pattern under a simple
linear regression model setting:

\begin{align} \label{eq:heter-model}
\boldsymbol{y} = 1 + \boldsymbol{x} + \boldsymbol{\varepsilon},\\
\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}, 1 + 2 - |a| b (\boldsymbol{x} - a)^2 \boldsymbol{I}), \\
\end{align}

where \(\boldsymbol{y}\), \(\boldsymbol{x}\),
\(\boldsymbol{\varepsilon}\) are vector of size \(n\).

The null regression model used to fit the realizations generated by the
above model is formulated exactly the same as the model used in the
first experiment:

\begin{equation} 
\boldsymbol{y} = \beta_0 + \beta_1 \boldsymbol{x} + \boldsymbol{u},
\end{equation}

Model misspecification presents since the assumption about the constant
error variance is violated.

Experiment data were simulated using three different shapes (\(a\) = -1,
0, 1), five different values of \(b\) (b = 0.25, 1, 4, 16, 64), three
different sample sizes (n = 50, 100, 300) and four different
distribution of \(X_{raw}\): (1) \(U(-1, 1)\), (2) \(N(0, 0.3^2)\), (3)
\(lognormal(0, 0.6^2)/3\) and (4) \(u\{1, 5\}\). A summary of the
parameters used in this experiment is given in Table
\ref{tab:parameter-table}.

The values of \(a\) was chosen so that different shapes of
heteroskedasticity were included in the residual plot. These include
triangle shape, butterfly shape and inverse triangle shape.

\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/different-shape-of-heter-1}

The range of \(b\), which is a factor controlling the strength of the
signal, was chosen so that different difficulty levels of lineups were
generated, and therefore, the estimated power curve would be smooth and
continuous.

\hypertarget{analysis-and-results}{%
\section{Analysis and results}\label{analysis-and-results}}

\hypertarget{data-overview}{%
\subsection{Data overview}\label{data-overview}}

How many people? How many lineups?

Subjects recruited from Prolific received a fixed payment for
participating in the experiment. However, some subjects will try to
maximize their earnings for minimum effort. During the review of
submissions, if we found a subject objectively demonstrated clear
low-effort throughout the experiment, i.e., failed all attention checks,
we rejected the submission. The rejected submissions will be removed
immediately, and Prolific will automatically recruit another subject as
substitution. Therefore, we only paid for approved submissions and no
further data screening procedure needed to be applied on the collected
data.

A subject was allowed to select zero plots for a lineup if there was no
visible difference between plots, but the simulated visual p-value given
in Equation (\ref{eq:p-value-multi}) will effectively drop the subject
from the simulation for this case, leading to inaccurate estimation of
\(p\)-value. Therefore, we treated this case as making one selection but
failing to identify the actual data plot so that Equation
(\ref{eq:p-value-multi}) can be applied correctly.

In overall, there were a total of 3200 lineup evaluations made by 160
subjects in both experiment I and experiment II respectively, where 320
lineup evaluations were attention checks and were not used in the
following analysis.

The collated dataset is provided in \texttt{polynomials} and
\texttt{heter} of the \texttt{visage} \texttt{R} package.

\hypertarget{section}{%
\subsection{}\label{section}}

\hypertarget{nonlinearity-1}{%
\subsection{Nonlinearity}\label{nonlinearity-1}}

\begin{itemize}
\tightlist
\item
  Power curve overall
\item
  Four lineups also shown, selected from ones close to the visual power
  curve from the uniform treatment
\item
  Compare visual power against different factors
\end{itemize}

\hypertarget{heteroskedasticity-2}{%
\subsection{Heteroskedasticity}\label{heteroskedasticity-2}}

\begin{itemize}
\tightlist
\item
  Power curve overall
\item
  Four lineups also shown, selected from ones close to the visual power
  curve from the uniform treatment
\item
  Compare visual power against different factors
\end{itemize}

\hypertarget{demographic-summary}{%
\subsection{Demographic summary}\label{demographic-summary}}

Table \ref{tab:demographic-table} tabulates the number of subjects,
preferred pronouns, education backgrounds, age groups, and previous
experience in visual experiment. Figure
\ref{fig:demographic-summary-plot} visualizes the marginal distribution
of each of the category. The collated data was a balanced sample among
male and female. Most of the ages of subject were between 18 to 39,
while many of them were between 18 to 24. Very few subjects were awarded
degree higher than Bachelor Degree. Around 40\% of subjects had previous
experience in visual experiment.

\caption{\label{tab:demographic-table}Summary of demographic information}
\centering
\begin{tabular}[t]{llllr}
\toprule
Preferred pronoun & Education & Age group & Previous experience & Subject\\
\midrule
She & Diploma and Bachelor Degree & 18-24 & Yes & 26\\
She & Diploma and Bachelor Degree & 18-24 & No & 24\\
He & Diploma and Bachelor Degree & 18-24 & No & 22\\
He & High School or below & 18-24 & No & 22\\
She & Diploma and Bachelor Degree & 25-39 & No & 21\\
He & Diploma and Bachelor Degree & 18-24 & Yes & 19\\
He & Diploma and Bachelor Degree & 25-39 & No & 19\\
She & High School or below & 18-24 & No & 19\\
He & High School or below & 18-24 & Yes & 15\\
She & Diploma and Bachelor Degree & 25-39 & Yes & 15\\
He & Diploma and Bachelor Degree & 25-39 & Yes & 12\\
She & High School or below & 18-24 & Yes & 10\\
He & Masters Degree & 25-39 & No & 8\\
He & Masters Degree & 25-39 & Yes & 8\\
He & High School or below & 25-39 & No & 7\\
She & Masters Degree & 25-39 & Yes & 7\\
She & High School or below & 25-39 & No & 6\\
He & High School or below & 25-39 & Yes & 5\\
She & Honours Degree & 25-39 & No & 5\\
She & Honours Degree & 25-39 & Yes & 4\\
She & Masters Degree & 25-39 & No & 4\\
He & Honours Degree & 25-39 & Yes & 3\\
She & Honours Degree & 18-24 & No & 3\\
He & Diploma and Bachelor Degree & 40-54 & Yes & 2\\
He & Diploma and Bachelor Degree & 55-64 & No & 2\\
He & Honours Degree & 25-39 & No & 2\\
He & Masters Degree & 40-54 & No & 2\\
She & Diploma and Bachelor Degree & 40-54 & No & 2\\
She & High School or below & 25-39 & Yes & 2\\
She & Masters Degree & 18-24 & No & 2\\
They & Diploma and Bachelor Degree & 18-24 & No & 2\\
They & Diploma and Bachelor Degree & 25-39 & No & 2\\
He & Diploma and Bachelor Degree & 40-54 & No & 1\\
He & High School or below & 40-54 & Yes & 1\\
He & High School or below & 55-64 & No & 1\\
He & Honours Degree & 18-24 & No & 1\\
He & Honours Degree & 18-24 & Yes & 1\\
He & Honours Degree & 40-54 & Yes & 1\\
He & Masters Degree & 40-54 & Yes & 1\\
He & Masters Degree & 55-64 & No & 1\\
Other & Diploma and Bachelor Degree & 18-24 & No & 1\\
Other & High School or below & 25-39 & No & 1\\
She & Diploma and Bachelor Degree & 40-54 & Yes & 1\\
She & High School or below & 40-54 & No & 1\\
She & High School or below & 55-64 & Yes & 1\\
She & Honours Degree & 40-54 & Yes & 1\\
She & Masters Degree & 40-54 & No & 1\\
They & High School or below & 18-24 & No & 1\\
They & High School or below & 18-24 & Yes & 1\\
They & High School or below & 25-39 & No & 1\\
\bottomrule
\end{tabular}
\end{table}

\begin{figure}
\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/demographic-summary-plot-1} \caption{Summary of demographic information. \label{fig:demographic-summary-plot}}\label{fig:demographic-summary-plot}
\end{figure}

\hypertarget{model-fitting}{%
\subsection{Model fitting}\label{model-fitting}}

For each parameter combination, effect \(E\) is derived from the
Kullback-Leibler divergence (see {[}appendix ref here{]}) formulated as:

\begin{equation} \label{eq:effect-size-ex1}
E = \frac{1}{2\sigma^2}\boldsymbol{X}_b'\boldsymbol{R}_a'(diag(\boldsymbol{R}_a))^{-1}\boldsymbol{R}_a\boldsymbol{X}_b,
\end{equation}

where \(diag(.)\) is the diagonal matrix constructed from the diagonal
elements of \(\boldsymbol{R}_a\).

The logistic regression is fit using \(log_e(E)\) as the only fixed
effect covariate for the power of visual test formulated as:

\begin{equation} \label{eq:logistic-regression-1-1}
Pr(\text{reject}~H_0|H_1,E) = \Lambda(\beta_0 + \beta_1 log_e(\boldsymbol{E})),
\end{equation}

where \(\Lambda(.)\) is the standard logistic function given as
\(\Lambda(z) = exp(z)/(1+exp(z))\).

To study various factors contributing to the power of the visual test,
the same logistic regression model is fit on different subsets of the
collated data grouped by levels of factors. This includes
{[}expansion{]}.

Table {[}table ref here{]} shows the parameter estimates of the logistic
regressions. {[}Discussion about the numeric estimates here{]}

\hypertarget{power-comparison}{%
\subsection{Power comparison}\label{power-comparison}}

\hypertarget{experiment-i}{%
\subsubsection{Experiment I}\label{experiment-i}}

Figure \ref{fig:power-overview} shows an overview of estimated power of
visual test against natural logarithm of the effect with comparison to
the power of an exact test - F-test, and the power of two other
residual-based conventional tests commonly used in regression
diagnostics but for testing other departures from the model assumptions.
In overall, the power of all four tests increases as the effect becomes
larger. The power curve of F-test climbs aggressively from 25\% to
around 90\% as \(log_e(E)\) increases from 0 to 2, while others respond
inactively to the change of effect and remain lower than 25\% throughout
the period, showing that as an exact test, the F-test is relatively more
sensitive to the type of model defects that being considered. The power
of visual test arises steadily and nearly linearly to around 90\% as
\(log_e(E)\) increases from 2 to 5, suggesting that the effect starts to
make noticeable impact on the degree of the presence of the designed
visual features. Other two inappropriate conventional tests shows
improvement at the same time but at a lower rate. This coincides the
point made by \citet{cook1982residuals} mentioned in
\ref{hypothesis-testing} that residual-based tests for a specific type
of model defect are sensitive to other types of model defects. At
\(log_e(E) = 6\), the power curve of F-test reaches almost 100\%
followed by the visual test by a small margin. The power of
Breusch--Pagan test and Shapiro--Wilk test reach around 75\% and 63\%
respectively.

What truly impress us is the huge difference between the estimated power
of visual test and the estimated power of F-test. The margin is largest
at around \(log_e(E) = 2\). An example lineup is included in Figure
\ref{fig:power-overview} where none of subjects detect the actual data
plot positioned at panel 14. It demonstrates that at this level of
difficulty, the designed visual feature is rarely visible, making the
actual data plot indistinguishable from residual plots simulated from
the assumed model. From a communication perspective, given the fact that
the visual difference is unperceivable, the argument that non-linearity
present in the fitted model is less convincing to the public even though
it is true. At around \(log_e(E) = 3\), the margin gets smaller as the
chance of identifying the actual data plot becomes larger. At this level
of difficulty, the designed visual features are usually detectable but
it may not stand out from the lineup as other null plots may happen to
include outliers or visual patterns that are are considered to be more
attractive by human, and thus recognized as the most different plot.
Without knowing the designed visual features beforehand, it is actually
hard to identify the actual data plot by pure image comparison. The
corresponding example lineup for \(log_e(E) = 3\) shown in Figure
\ref{fig:power-overview} has the actual data plot positioned at panel
20, where two out of five subjects detect it. It can be observed that a
M-shape is presented in plot 20, but the signal is not strong enough to
attract all five subjects, resulting in a visual p-value sightly above
the desired significance level \(\alpha = 0.05\). At \(log_e(E) = 4\)
and \(log_e(E) = 6\), the designed visual features become much clear and
attractive, leading to a high percentage of rejection of the null
hypothesis. Figure \ref{fig:power-overview} gives example lineups of
such cases.

\begin{figure}
\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/power-overview-1} \caption{Power overview. \label{fig:power-overview}}\label{fig:power-overview}
\end{figure}

\hypertarget{distributions-of-regressor}{%
\subsubsection{Distributions of
regressor}\label{distributions-of-regressor}}

The impact of the distribution of \(X_raw\) on the power is shown in
Figure \ref{fig:dist-power}. The power curve of F-test is stable across
different distributions, while the visual test has a steeper power curve
for normal and uniform distribution. BP-test performs worse for discrete
uniform distribution and uniform distribution but has relatively high
power for normal distribution. SW-test outperforms BP-test for discrete
uniform distribution but remains as the worst test for other
distributions. The results indicate those inappropriate residual-based
tests are sensitive to the distribution of the regressor.

\begin{figure}
\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/dist-power-1} \caption{\label{dist-power}}\label{fig:dist-power}
\end{figure}

\hypertarget{shpaes-of-non-linearity}{%
\paragraph{Shpaes of non-linearity}\label{shpaes-of-non-linearity}}

Figure \ref{fig:shape-power} illustrates the change of power under
different shape of non-linearity. Similar to the power curves shown in
\ref{fig:dist-power}, F-test is stable under different shapes. The power
curve of visual test also behaves similarly across different shapes.
What vary are the power curve of BP-test and SW-test. For Triple-U
shape, both BP-test and SW-test are insensitive to the change of the
effect. And for W shape, both tests have almost identical power curves.
It can be observed that both tests performs the best for U-shape.

\begin{figure}
\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/shape-power-1} \caption{\label{shape-power}}\label{fig:shape-power}
\end{figure}

\hypertarget{reasons-for-making-selections}{%
\paragraph{Reasons for making
selections}\label{reasons-for-making-selections}}

\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-13-1}

\hypertarget{experiment-ii}{%
\subsubsection{Experiment II}\label{experiment-ii}}

\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-16-1}

\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-17-1}

\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-18-1}

\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-19-1}

\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-20-1}

>>>>

With our knowledge, the effectiveness of visual test relative to the
equivalent conventional test in regression diagnostics has not been
validated. Particularly, its ability to detect non-linearity and
heteroskedasticity compared to model structural test and Breusch--Pagan
test. Meanwhile, as mentioned in Section \ref{introduction} section, we
keen to understand the differences between conventional hypothesis
testing and graphical diagnostics in the application of linear
regression diagnostics and why conventional tests are not preferred by
data analysts in linear regression diagnostics. Attempting to find the
answer to these questions, three experiments were conducted.

The experiment I has ideal scenario for conventional testing, where the
visual test is not expected to outperform the conventional test. The
experiment II is a scenario where the conventional test is an
approximate test, in which the visual test may have a chance to match
the performance of the conventional test. The experiment III is designed
for collecting human responses to lineup with only good residual plots
such that the parameter \(\alpha\) in Equation
\ref{eq:pvalue-beta-binomial} can be estimated.

Subjects for all three experiments were recruited from an crowdsourcing
platform called Prolific {[}ref here{]}. Prescreening procedure was
applied during the recruitment, subjects were required to be fluent in
English, with \(98\%\) minimum approval rate in other studies and 10
minimum submissions.

@@

\hypertarget{non-linearity}{%
\subsubsection{Non-linearity}\label{non-linearity}}

@@

Since \(z = O(x^j)\), for \(j > 1\), \(z\) is a higher order term leaves
out by the null regression causing model misspecification.

@@
